Strong coupling between scales in a multi-scalar model of urban dynamics

### 2021-06-25 Docs/Papers/Multiscale/paper/paper.tex
.
bib
revisions

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0003-0768-9480} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad CASA, University College London\\
$^{2}$ \quad Institut des Syst{\`e}mes Complexes Paris-Ile-de-France, CNRS\\
$^{3}$ \quad UMR G{\'e}ographie-cit{\'e}s, CNRS}

@@

@@

@@

% ! reformulate - this is the CCS abstract

@@

To couple the scales, we assume that (i) macroscopic performance of a city will influence the choices made by planers in terms of land-use, and thus parameters of the morphogenesis mode locally - empirical evidence of such links have been suggested in the literature \cite{joy2015toronto} and we stylise them here; (ii) urban form influences the inner workings of a city through congestion and the fostering of exchanges between agents (individuals, firms), and finally the global insertion of the city - thus spatial interaction parameters are updated as a function of local urban form. This last link is the most % discutable % TODO slippery?
 in terms of empirical evidence and implies many other dimensions than population only. We however work on a stylised explicative model which must keep a simple structure but on which a systematic knowledge of model behavior can be developed. We show in Fig.~\ref{fig:fig1} a schematic description of model structure.

\begin{figure}[H]
%  figure: more details?
 \includegraphics[width=\linewidth]{figures/multiscale_morph.pdf}
 \caption{Schematic representation of model architecture. Urban areas with their own urban form are inserted into the macroscopic spatial interaction model, and strong coupling is achieved through top-down and bottom-up feedbacks.\label{fig:fig1}}
\end{figure}

@@

The simulation model operates iteratively, and at each time step goes sequentially through the following steps:
\begin{enumerate}
	\item populations of urban areas are evolved at the macroscopic scale with the spatial interaction model (Equations \ref{eq:eq1} to \ref{eq:eq3} below);
	\item for each urban area, parameters of the moprhogenesis model are updated given parameters capturing typical scenarios: transit-oriented development or sprawl for the diffusion parameter, metropolisation or uniformisation for the aggregation parameter (Equations \ref{eq:eq4} to \ref{eq:eq6});
	\item local urban form are evolved conditionally to a fixed population increase and following the mescoscopic parameters;
	\item congestions and interactions within areas are synthesised following a cost function, which is used to update the macroscopic interaction parameters (Equations \ref{eq:eq7} and \ref{eq:eq8}).
\end{enumerate}

@@

\begin{equation}
	\label{eq:eq5}
	Z_i = \sum_j \frac{P_j}{\sum_k P_k} \cdot \exp( - d_{ij} / d_i)
\end{equation}

@@

\begin{equation}
\label{eq:eq6}
		U_i = \sum_{kl} \left( \frac{P_k P_l}{P^2} \cdot \frac{1}{d_{kl}} - \lambda \left(\frac{P_k P_l}{P^2} \cdot \frac{1}{d_{kl}}\right)^2 \right)
\end{equation}

@@

\begin{equation}
\label{eq:eq7}
	d_i (t+1) = d_i (t) \left( 1 + \delta d \cdot \frac{U_i}{\max_k \left|U_k\right|} \right)
\end{equation}

@@

The Table~\ref{tab:parameters} summarizes model parameters. For clarity, we do not include parameters linked to synthetic model setup, and distinguish them by the scale to which they are attached.

\begin{table}[H]
	\caption{Summary of model parameters\label{tab:parameters}}
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
		Type & Parameter & Process & Range \\\hline
		\multirow{4}{*}{Macro} & $g_i = g_0$ & Endogenous growth & $\left[0 ; 0.5\right]$ \\
		& $w_i = w_G$ & Interactions weight & $\left[0 ; 0.1\right]$ \\
		& $\gamma_i = \gamma_G$ & Interactions hierarchy & $\left[0 ; 5.0\right]$ \\
		& $d_i$ & Interactions decay & $\left[0 ; 1000\right]$ \\ \hline
		\multirow{4}{*}{Meso} & $\alpha_i$ & Aggregation & $\left[0 ; 5\right]$ \\
		& $\beta_i$ & Diffusion & $\left[0 ; 0.5\right]$ \\
		& $t_m$ & Urban growth speed & $\left[0 ; 20\right]$ \\
		& $n_d$ & Diffusion & $\left[0 ; 5\right]$ \\ \hline
		\multirow{4}{*}{Multiscale} & $\delta\alpha$ & Downward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\delta\beta$ & Downward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\delta d$ & Upward feedback & $\left[-0.5 ; 0.5\right]$ \\\hline
		& $\lambda$ & Cost of congestion & $\left[0 ; 2\right]$ \\\hline
	\end{tabular}
\end{table}

@@

@@

Model implementation is done under some performance constraints: $N$ mesoscopic morphogenesis models must be simulated in parallel at each macroscopic time step. On the contrary, macroscopic interactions are efficient to implement as they are based on matrices computations. The model is implemented in \texttt{scala} and integrated into a broader library for spatial sensitivity analysis and dynamical models for settlement systems, namely the \texttt{spatialdata} library \cite{raimbault2020scala}. This library in particular already includes implementations of the models used in \cite{raimbault2020indirect} and \cite{raimbault2018calibration}, but also of other urban interaction models such as \cite{favaro2011gibrat}. The relative large number of parameters and output indicators imposes the use of appropriate model exploration methods to get a grasp of model behavior. Therefore, the parameter space is explored with the OpenMOLE model exploration software \cite{reuillon2013openmole}. Standalone source code for the model is available on an open git repository at \url{https://github.com/JusteRaimbault/UrbanGrowth-model}, while the repository for results is at \url{https://github.com/JusteRaimbault/UrbanGrowth}. Simulation data used in this paper is available on the dataverse repository at \url{https://doi.org/10.7910/DVN/IRHMQK}.

@@

@@

>>>>

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0003-0768-9480} % Add \orcidA{} behind the author's name

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Center for Advanced Spatial Analysis, University College London\\
$^{2}$ \quad Institut des Syst{\`e}mes Complexes Paris-Ile-de-France, CNRS\\
$^{3}$ \quad UMR G{\'e}ographie-cit{\'e}s, CNRS}

@@

\cite{lockerby2015asynchronous} physics multiscale simulations
+ multiscale traffic models (Nancy?)

\cite{vinci2021population} : link macro migration - meso structure (polycentrism)

def of scale \cite{atkinson2000spatial}

\cite{jantz2005analysis}

@@

% TODO emphasize more the research problem

@@

% TODO ! reformulate - this is the CCS abstract

@@

To couple the scales, we assume that (i) macroscopic performance of a city will influence the choices made by planers in terms of land-use, and thus parameters of the morphogenesis mode locally - empirical evidence of such links have been suggested in the literature \cite{joy2015toronto} and we stylise them here; (ii) urban form influences the inner workings of a city through congestion and the fostering of exchanges between agents (individuals, firms), and finally the global insertion of the city - thus spatial interaction parameters are updated as a function of local urban form. This last link is the most questionable assumption in terms of empirical evidence and implies many other dimensions than population only. We however work on a stylised explicative model which must keep a simple structure but on which a systematic knowledge of model behavior can be developed. We show in Fig.~\ref{fig:fig1} a schematic description of model structure.

\begin{figure}[H]
% TODO figure: more details?
 \includegraphics[width=\linewidth]{figures/multiscale_morph.pdf}
 \caption{Schematic representation of the model architecture. Urban areas with their own urban form are inserted into the macroscopic spatial interaction model, and strong coupling is achieved through top-down and bottom-up feedbacks.\label{fig:fig1}}
\end{figure}

@@

The simulation model operates iteratively, and at each time step goes sequentially through the following steps:
\begin{enumerate}
	\item populations of urban areas are evolved at the macroscopic scale with the spatial interaction model (Equations \ref{eq:eq1} to \ref{eq:eq3} below);
	\item for each urban area, parameters of the morphogenesis model are updated given parameters capturing typical scenarios: transit-oriented development or sprawl for the diffusion parameter, metropolisation or uniformisation for the aggregation parameter (Equations \ref{eq:eq4} to \ref{eq:eq6});
	\item local urban form are evolved conditionally to a fixed population increase and following the mesoscopic parameters;
	\item congestions and interactions within areas are synthesised following a cost function, which is used to update the macroscopic interaction parameters (Equations \ref{eq:eq7} and \ref{eq:eq8}).
\end{enumerate}

@@

\begin{equation}
	\label{eq:eq6}
	Z_i = \sum_j \frac{P_j}{\sum_k P_k} \cdot \exp( - d_{ij} / d_i)
\end{equation}

@@

\begin{equation}
\label{eq:eq7}
		U_i = \sum_{kl} \left( \frac{P_k P_l}{P^2} \cdot \frac{1}{d_{kl}} - \lambda \left(\frac{P_k P_l}{P^2} \cdot \frac{1}{d_{kl}}\right)^2 \right)
\end{equation}

@@

\begin{equation}
\label{eq:eq8}
	d_i (t+1) = d_i (t) \left( 1 + \delta d \cdot \frac{U_i}{\max_k \left|U_k\right|} \right)
\end{equation}

@@

The Table~\ref{tab:parameters} summarises model parameters. For clarity, we do not include parameters linked to synthetic model setup, and distinguish them by the scale to which they are attached.
% TODO ref table beginning section

\begin{table}[H]
	\caption{Summary of model parameters\label{tab:parameters}}
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
		Type & Parameter & Process & Range \\\hline
		\multirow{4}{*}{Macro} & $g_i = g_0$ & Endogenous growth & $\left[0 ; 0.5\right]$ \\
		& $w_i = w_G$ & Interactions weight & $\left[0 ; 0.1\right]$ \\
		& $\gamma_i = \gamma_G$ & Interactions hierarchy & $\left[0 ; 5.0\right]$ \\
		& $d_i$ & Interactions decay & $\left[0 ; 1000\right]$ \\ \hline
		\multirow{4}{*}{Meso} & $\alpha_i$ & Aggregation & $\left[0 ; 5\right]$ \\
		& $\beta_i$ & Diffusion & $\left[0 ; 0.5\right]$ \\
		& $t_m$ & Urban growth speed & $\left[0 ; 20\right]$ \\
		& $n_d$ & Diffusion & $\left[0 ; 5\right]$ \\ \hline
		\multirow{4}{*}{Multiscale} & $\delta\alpha$ & Downward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\delta\beta$ & Downward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\delta d$ & Upward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\lambda$ & Cost of congestion & $\left[0 ; 2\right]$ \\\hline
	\end{tabular}
\end{table}

@@

% TODO extend indicator descritpion

@@

Model implementation is done under some performance constraints: $N$ mesoscopic morphogenesis models must be simulated in parallel at each macroscopic time step. On the contrary, macroscopic interactions are efficient to implement as they are based on matrices computations. The model is implemented in \texttt{scala} and integrated into a broader library for spatial sensitivity analysis and dynamical models for settlement systems, namely the \texttt{spatialdata} library \cite{raimbault2020scala}. This library in particular already includes implementations of the models used in \cite{raimbault2020indirect} and \cite{raimbault2018calibration}, but also of other urban interaction models such as \cite{favaro2011gibrat}. The relative large number of parameters and output indicators imposes the use of appropriate model exploration methods to get a grasp of model behavior. Therefore, the parameter space is explored with the OpenMOLE model exploration software \cite{reuillon2013openmole}. Standalone source code for the model is available on an open git repository as \cite{urbangrowthmodel}. An other open repository for the analysis of results is also available as \cite{urbangrowth}. Simulation data used in this paper is available on the dataverse repository as \cite{dataverse}.

@@

% TODO improve discussion - link with literature: related results, interpretation regarding previous studies

% TODO limitations of the work

@@

% TODO implication of results in a broad context
